---
title: "Consumer reports AVR data"
output: word_document
---

These summary statistics capture the number of Aortic valve replacement surgeries reported by each hospital between 2011 and 2014, along with the 30-day survival rate and percentage of patients with none of 5 major kinds of complications. (Complications include  five serious complications of heart-bypass surgery during their hospitalization: extended breathing support on a ventilator, an infection in the breastbone incision, kidney failure, a stroke, or a repeat operation for postoperative bleeding or other causes).

```{r, echo=FALSE}
setwd("C:/Users/steph/Documents/AVR/Consumer Report data/Data/raw_data")
data<-read.csv("Raw results (original download).csv")

#prints table of total patients, survival % and complication % incl min, max, mean, and median, as well as # of NAs
summary(data[,4:6])

#Removing NA values in order to plot
data_present<-data[complete.cases(data),]

#Creating new variable of No complication rate added to survival rate
Aggregate_score<-data_present$No.complication + data_present$Survival

```

Plotting the no-complication rate against total patients revealed a weak relationship between volume of operations completed.

```{r, echo=FALSE}
#Fitting a regression line for 30 day no-complication rate plotted against total patients
plot(data_present$No.complication, data_present$Total.Patients, pch = 16, col = "red", main = "No complication rate plotted against total patients", xlab = "No complication rate", ylab = "Total patients undergoing AVR 2011-2014")
abline(lm(data_present$Total.Patients ~ data_present$No.complication))
```

Plotting the survival rate against total patients also revealed a weak relationship. However, particularly with low-volume centers with low sample size (n<20), the statistical significance of the differences between centers is questionable.

```{r, echo=FALSE}

#Fitting a regression line for survival plotted against total patients
plot(data_present$Survival, data_present$Total.Patients, pch = 16, col = "blue", main = "Survival plotted against total patients", xlab = "survival (30 day)", ylab = "Total patients undergoing AVR 2011-2014")
abline(lm(data_present$Total.Patients ~ data_present$Survival))

```

The aggregate score was calculated by adding the two metrics for each hospital (no complication rate and survival rate).

```{r, echo=FALSE}
#Fitting a regression line for Aggregate score plotted against total patients
plot(Aggregate_score, data_present$Total.Patients, pch = 16, cex = 1.3,col = "green", main = "Aggregate score plotted against total patients", xlab = "Aggregate Score", ylab = "Total patients undergoing AVR 2011-2014")
abline(lm(data_present$Total.Patients ~ Aggregate_score))
```

```{r, echo=FALSE}
#Fitting a regression line for 30 day no-complication rate plotted against survival rate
plot(data_present$Survival, data_present$No.complication, pch = 16, col = "purple", main = "No complication rate plotted against survival rate", xlab = "No complication rate", ylab = "Survival rate")
abline(lm(data_present$Survival ~ data_present$No.complication))
```
